Clustering and Parameter Estimation for the Popularity Adjusted Block Model

### 2020-07-13 summary.Rmd
minor updates to writeup

\citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
to identify the cluster memberships given either an edge probability matrix $P$ 
or an adjacency matrix $A$. In the case that $P$ is known, the cluster 
memberships can be identified exactly (up to permutation). A similar procedure 
can be applied if $P$ is unknown and we have an observation $A$, but the 
theoretical guarantees of this method applied to the PABM are unknown. In 
particular, the method requires spherical Gaussian noise. The authors of this 
paper then use point estimators for $\{\lambda^{(kl)}\}$ with the results of 
SSC.

@@

**Definition 1**
Let $G = (V, E)$ be an undirected, unweighted random graph with corresponding 
affinity matrix $A \in \{0, 1\}^{n \times n}$. Then $A$ is a random matrix 
with corresponding edge probability matrix $P$ such that 
$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ 
($A_{ji} = A_{ij}$ and $A_{ii} = 0$). Let there exist $K$ underlying 
communities in $G$, and let $n_{k}$ be the size of the $k$^th^ community in 
$G$ such that $\sum_{k=1}^K n_k = n$.

@@

**Definition 2**
Let $X \in \mathbb{R}^{n \times d}$ be latent positions of the vertices of a 
graph $G$. $X$ consists of row vectors $x_i^\top$. Let 
$A \in \{0, 1\}^{n \times n}$ be the corresponding affinity matrix.

@@

>>>>

\citet*{noroozi2019estimation} proposed using sparse subspace clustering (SSC) 
to identify the cluster memberships given either an edge probability matrix $P$ 
or an adjacency matrix $A$. In the case that $P$ is known, the cluster 
memberships can be identified exactly (up to permutation). A similar procedure 
can be applied if $P$ is unknown and we have an observation $A$, but the 
theoretical guarantees of this method applied to the PABM are unknown. In 
particular, the method requires spherical Gaussian noise. The authors of this 
paper then use point estimators for $\{\lambda^{(kl)}\}$ using cluster labels 
obtained via SSC.

@@

**Definition 1**.
Let $G = (V, E)$ be an undirected, unweighted random graph with corresponding 
affinity matrix $A \in \{0, 1\}^{n \times n}$. Then $A$ is a random matrix 
with corresponding edge probability matrix $P$ such that 
$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i > j$ 
($A_{ji} = A_{ij}$ and $A_{ii} = 0$). Let there exist $K$ underlying 
communities in $G$, and let $n_{k}$ be the size of the $k$^th^ community in 
$G$ such that $\sum_{k=1}^K n_k = n$.

@@

**Definition 2**.
Let $X \in \mathbb{R}^{n \times d}$ be latent positions of the vertices of a 
graph $G$. $X$ consists of row vectors $x_i^\top$. Let 
$A \in \{0, 1\}^{n \times n}$ be the corresponding affinity matrix.

@@

**Remark**. 
Like the "ungeneralized" random dot product graph, the latent positions of a 
generalized random dot product graph (RDPG) are not unique. More specifically, 
if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
$Q \in \{R : R^\top I_{pq} R = I_{pq} \}$, 
$(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
x_i^\top I_{pq} x_j = P_{ij}$ (this is the indefinite orthogonal group, 
$\mathbb{O}(p, q)$).
Unlike in the RDPG case, transforming the latent positions by multiplication 
with $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or 
distances.
